\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-08-19 gwnnDraft02.tex
Fixed red comments

Immediately from Eq. (\ref{eqn:nnEffLag}), we can see that the matrix element for decays into standard sectors is inversely proportional to that sectors higgs mass, $\cM_{m_H^2 < 0} \sim 1/m_{h_i}$. The Feynman diagram for this process is presented in Fig. \ref{fig:reheatonDecays}. The loop decay of $\phi \rightarrow \gamma\gamma$ is always sub-leading and can be neglected. It should be noted that as one goes to sectors with larger and larger vevs, the increasing mass of the fermions ($m_f \sim v_i \sim v_{us}\sqrt{i}$) eventually leads to situations where the decay to two on-shell bottom or charm quarks is kinematically forbidden, $m_\phi < 2 m_q$. \textcolor{green}{Put an equation sketching how $m_f$ scales with $m_h$} For sectors where this kinematic threshold is passed for charm quarks, contributions to cosmological observables can be safely ignored. All in all, we end up with a decay width that scales as $\Gamma_{m_H^2<0} \sim 1/m_h^2$. Since we can expect energy density to be proportional to the decay width, $\frac{\rho_i}{\rho_{SM}} \approx \frac{\Gamma_i}{\Gamma_{SM}}$, this indicates that energy density of standard sectors falls:
\begin{equation}\label{eqn:edSS}
\rho_i \sim \frac{\rho_{SM}}{i}.
\end{equation}

@@

The critical index demonstrates that even in the most extreme cases, where first order phase transitions occur with the up quark an order of magnitude less than $\Lambda_{QCD}$, the critical index is still $\sim 1000$. Looking back to the energy density scaling in Eq.~(\ref{eqn:edSS})\textcolor{green}{refer to a particular equation} we see that, within the context of $N$naturalness, the energy density of these sectors is at most $10^{-4}$ \textcolor{red}{why not $10^{-3}$?} that of our sector. As a result, \textcolor{red}{As we will show below?} the gravitational waves generated by the standard large $i$ sectors in $N$naturalness will not generate detectable signatures.

@@

The masses obtained through the phase transition can be approximated through the use of a generalization of the Gell-Mann--Oakes--Renner relation \textcolor{green}{Also cite Gell-Mann et al} \cite{Gell-Mann,Schwartz:2013pla},
\begin{equation}\label{eqn:gmor}
m^2_{\pi} = \frac{V^3}{F^2_{\pi}}(m_u + m_d),
\end{equation}
where $V \sim \Lambda_{QCD}$, $F_\pi$ is the pion decay constant. So, for mesons in exotic sector $i$:
\begin{equation}
m_{\pi}^i \sim \sqrt{\frac{m_1^i+m_2^i}{m_u + m_d}} m_{\pi}.
\end{equation}
\textcolor{red}{What are $m_a^i$? Is this the right formula,isn't $V$ and $f_\pi$ different from the SM to the $i$th exotic sector? Also, how do you compute $\Delta m$ when you have a different set of particles inside (pions) as outside (quarks) the bubble?  } Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Thus, all exotic sectors have strong first order phase transitions that lead to runaway bubble walls.

@@

The original $N$naturalness paper focused on the (stronger) $N_{eff}$ constraints from the additional standard sectors and ignored the exotic sectors altogether \cite{Arkani-Hamed:2016rle}; in contrast, we take the opposite approach. Since we're more interested in a more general scenario with multiple dark phase transitions, we focus in on whether zero vev sectors can dodge these $N_{eff}$ bounds.

Within the framework of SM-like particle content (save the higgs mass parameter), our exotic sectors feature at most 106.75 effective degrees of freedom at high energies: 72 from quarks,  12 from charged leptons, 6 from neutrinos, 16 from gluons, 2 from photons, 6 from W bosons (no longitudinal modes due to a no symmetry breaking), 4 from the higgs sector (again, no symmetry breaking leads to 3 additional modes), for a total of 118 degress of freedom. Scaling the fermionic contributions by $7/8$ gives us 106.75 effective relativistic degrees of freedom.  After the higgses freeze out, 4 degrees of freedom are lost; when the QCD phase transition occurs 35 bosonic degrees of freedom are gained while 72 fermionic are lost, leaving 56.75 effective degrees of freedom. Since all additional sectors are colder than our own \textcolor{yellow}{Aren't they always colder than our own? Yes, they are.} by the time photon decoupling occurs ($T_\gamma \sim 0.39 \,\ev$), the pions of all of these sectors will, in accordance with Eq.~(\ref{eqn:gmor}), have a much higher mass than the sector temperature and thus will have frozen out long before --- leaving at most $11.25$ effective degrees of freedom per sector \textcolor{red}{From photons and neutrinos? What about $\Delta N_{eff}$ constraints from BBN? This would be a much higher (SM) temperature.} .

>>>>

Immediately from Eq. (\ref{eqn:nnEffLag}), we can see that the matrix element for decays into standard sectors is inversely proportional to that sectors higgs mass, $\cM_{m_H^2 < 0} \sim 1/m_{h_i}$. The Feynman diagram for this process is presented in Fig. \ref{fig:reheatonDecays}. The loop decay of $\phi \rightarrow \gamma\gamma$ is always sub-leading and can be neglected. It should be noted that as one goes to sectors with larger and larger vevs, the increasing mass of the fermions ($m_f \sim v_i \sim v_{SM}\sqrt{i}$) eventually leads to situations where the decay to two on-shell bottom or charm quarks is kinematically forbidden, $m_\phi < 2 m_q$. \textcolor{green}{Put an equation sketching how $m_f$ scales with $m_h$} For sectors where this kinematic threshold is passed for charm quarks, contributions to cosmological observables can be safely ignored. All in all, we end up with a decay width that scales as $\Gamma_{m_H^2<0} \sim 1/m_h^2$. Since we can expect energy density to be proportional to the decay width, $\frac{\rho_i}{\rho_{SM}} \approx \frac{\Gamma_i}{\Gamma_{SM}}$, this indicates that energy density of standard sectors falls:
\begin{equation}\label{eqn:edSS}
\rho_i \sim r_{s}\frac{\rho_{SM}}{i}
\end{equation} 
with $r_s$ being the ratio of the energy density of the first additional standard sector over the energy density of our sector.

@@

The critical index demonstrates that even in the most extreme cases, where first order phase transitions occur with the up quark an order of magnitude less than $\Lambda_{QCD}$, the critical index is still $\sim 1000$. Looking back to the energy density scaling in Eq.~(\ref{eqn:edSS})\textcolor{green}{refer to a particular equation} we see that, within the context of $N$naturalness, the energy density of these sectors is at most $10^{-4}$ \textcolor{green}{why not $10^{-3}$?} that of our sector. As a result, \textcolor{green}{As we will show below?} the gravitational waves generated by the standard large $i$ sectors in $N$naturalness will not generate detectable signatures.

@@

The masses obtained through the phase transition can be approximated through the use of a generalization of the Gell-Mann--Oakes--Renner relation \textcolor{green}{Also cite Gell-Mann et al} \cite{Gell-Mann,Schwartz:2013pla},
\begin{equation}\label{eqn:gmor}
m^2_{\pi} = \frac{V^3}{F^2_{\pi}}(m_u + m_d),
\end{equation}
where $V \sim \Lambda_{QCD}$, $F_\pi$ is the pion decay constant. One expects that within a given sector $F_\pi \sim V \sim \Lambda_{QCD}$ \cite{Schwartz:2013pla} and as exotic sectors have $\Lambda_{ex} \sim 90 \,\mev$ while the SM features $\Lambda_{QCD} = (332\pm17)\,\mev$ \cite{PhysRevD.98.030001} we expect at most $\mathcal{O}(1)$ difference in the $\sqrt{\frac{V^3}{F_\pi^2}}$ coefficient between sectors. So, for mesons in exotic sector $i$:
\begin{equation}
m_{\pi}^i \sim \sqrt{\frac{m_a^i+m_b^i}{m_a + m_b}} m_{\pi}.
\end{equation}
Here, $a$ and $b$ denote the component quark flavours. It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$. \textcolor{green}{What are $m_a^i$? Is this the right formula,isn't $V$ and $f_\pi$ different from the SM to the $i$th exotic sector? Also, how do you compute $\Delta m$ when you have a different set of particles inside (pions) as outside (quarks) the bubble?  }

Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Thus, all exotic sectors have strong first order phase transitions that lead to runaway bubble walls.

@@

The original $N$naturalness paper focused on the (stronger) $N_{eff}$ constraints from the additional standard sectors and ignored the exotic sectors altogether \cite{Arkani-Hamed:2016rle}; in contrast, we take the opposite approach. Since we're more interested in a more general scenario with multiple dark phase transitions, we focus in on whether zero vev sectors can dodge these $N_{eff}$ bounds. Additionally, 
\begin{equation}
\frac{(\Delta N^i_{eff})_{CMB}}{(\Delta N^i_{eff})_{BBN}} \geq 1
\end{equation}
for any sector $i$ \cite{Arkani-Hamed:2016rle}. Because the constraints on $N_{eff}$ are stronger at photon decoupling than at BBN, we can focus purely on the constraints provided by the former \cite{Aghanim:2018eyx}.

Within the framework of SM-like particle content (save the higgs mass parameter), our exotic sectors feature at most 106.75 effective degrees of freedom at high energies: 72 from quarks,  12 from charged leptons, 6 from neutrinos, 16 from gluons, 2 from photons, 6 from W bosons (no longitudinal modes due to a no symmetry breaking), 4 from the higgs sector (again, no symmetry breaking leads to 3 additional modes), for a total of 118 degress of freedom. Scaling the fermionic contributions by $7/8$ gives us 106.75 effective relativistic degrees of freedom.  After the higgses freeze out, 4 degrees of freedom are lost; when the QCD phase transition occurs 35 bosonic degrees of freedom are gained while 72 fermionic are lost, leaving 56.75 effective degrees of freedom. Since all additional sectors are colder than our own \textcolor{yellow}{Aren't they always colder than our own? Yes, they are.} by the time photon decoupling occurs ($T_\gamma \sim 0.39 \,\ev$), the pions of all of these sectors will, in accordance with Eq.~(\ref{eqn:gmor}), have a much higher mass than the sector temperature and thus will have frozen out long before --- leaving at most $11.25$ effective degrees of freedom per sector \textcolor{green}{From photons and neutrinos? What about $\Delta N_{eff}$ constraints from BBN? This would be a much higher (SM) temperature.} .
